---
title: "Include datasets in your package"
subtitle: "What about data?"
code: "M13S02C09"
output: formation::chapitre
prerequis: ["M06-rbase"]
obj_ped: "savoir intégrer des données dans un packages"
slug: encapsuler-donnees-en
---
class: slide 


### Include datasets

In a package, it can be useful to include data.

+ To demonstrate the use of a function with a relevant example
+ To disseminate information

--

There are three types of datasets to include, thus three ways to include them into a package. 

1. A raw data file (xlsx, csv or other), **NOT** available to the end user, but accessible to the developers only, stored in `data-raw/` folder

--

2. An example dataset in `rda` format, to be loaded as is, available to the user, stored in the `data/` folder (such as `iris` or `mtcars` for example)
<!-- > If you specified `LazyData: true` in your `DESCRIPTION`, these datasets will be loaded lazily, i.e., they will only be available if the user explicitly loads them.  -->

--

3. A data file (xlsx, csv or other) not transformed into `rda`, available to the end user, stored in the `inst/` folder


???

Let's see how and why

---
class: slide 

### 1. Include datasets in `data-raw/`
 
#### Internal dataset, accessible to developers only
 
- Developers store raw datasets to keep the source of examples close to them
- Users can not access them, there are not installed

--

*Steps:*  

- Create folder `data-raw/` at the root of the package using: `usethis::use_data_raw()`
- Insert raw datasets inside
- Use R/Rmd scripts to prepare a cleaner / smaller dataset for future examples  

--

*Information:*  

- `data-raw` is not installed with the package
  + It is not accessible to the user
- Example: <https://github.com/ThinkR-open/prenoms/tree/master/data-raw>

???

- Prepare data that you will need for your reproducible examples
- Keep the original source of your datasets in a safe place
- Note that with {fusen}, you can realise the preparation steps in a `development` chunk.

---
class: slide 

### 2. Include datasets in `data/`

#### Exported dataset, accessible to the user

- Developers includes a dataset in the package using: `usethis::use_data(my_dataset)` 
- Users load the dataset named `my_dataset` using: `data(my_dataset)`

--

*Steps:*  

- Create `data-raw/` to prepare your dataset using: `usethis::use_data_raw()`
- Prepare your dataset as needed for reproducible examples
- Store `my_dataset` as internal data

```{r eval = FALSE}
# Read some raw data 
# my_data_to_clean <- readr::read_csv("my_raw_data.csv")

# Or use existing dataset like `iris`
my_dataset <- dplyr::slice_sample(iris, prop = 0.2)
usethis::use_data(my_dataset, overwrite = TRUE) 
```

- `overwrite = TRUE` overwrite the dataset if already exists, as for an update
- Have a look at the content of newly created folder `data/`

???

- We use the directory `data-raw` previously created to prepare dataset
- `use_data()` stores the reproducible example at the right place in the right format

---
class: slide 

### 2. Include datasets in `data/`

#### Exported dataset, accessible to the user

- Developers includes a dataset in the package using: `usethis::use_data(my_dataset)` 
- Users load the dataset named `my_dataset` using: `data(my_dataset)`


*Information:*  

- `my_dataset` is accessible to the user after package installation
```{r, eval=FALSE}
library(mypackage)
data(my_dataset)
```
- `my_dataset` is stored as `.rda` file, only readable by R, similar to `.RData` files

???

Package needs to be installed to access the dataset

---
class: slide 

### 2. Include datasets in `data/`

#### Exported dataset, accessible to the user

- Exported datasets need to be documented, like functions

--

*Steps:*  

- Create a dedicated script in "R/" folder using `usethis::use_r("my_dataset")`
- Fill in the 'roxygen2' documentation to describe the dataset

```{r eval=FALSE}
#' Prices of 50,000 round cut diamonds.
#'
#' A dataset containing the prices and other attributes of almost 54,000
#' diamonds.
#'
#' @format A data frame with 53940 rows and 10 variables:
#' \describe{
#'   \item{price}{price, in US dollars}
#'   \item{carat}{weight of the diamond, in carats}
#'   ...
#' }
#' @source \url{http://www.diamondse.info/}
"diamonds"
```

???

Documentation is always the key for a correct package. 
Check will remind you of missing data documentation.


---
class: slide 

### 2. Include datasets in `data/`

#### Exported dataset, accessible to the user

- Exported datasets need to be documented, like functions

*Information:*  

- Documentation requires at least `@format` and `@source`
- We do not `@export` a dataset

---
class: slide 

### 3. Include datasets in `inst/`

#### Raw dataset, accessible to the user

- Developers includes a dataset in the `inst/` folder directly
- Users find the dataset using `system.file("my_dataset.csv", package = "mypackage")`

--

*Steps*

- Create `inst/` folder using: `dir.create(here::here("inst"))`
- Add a dataset inside
- Install the package
- Users can access the path of the dataset and read the file as usual

```{r, eval=FALSE}
# Store "my_dataset.csv" in "inst/" folder
the_data_path <- system.file("my_dataset.csv", package = "mypackage")
the_data <- readr::read_csv(the_data_path)
```

--

*Information*

- Any type of data is allowed
- Use the path for your reproducible examples
- Organise the content of `inst/` as your want

???

- system.file() does not read the dataset, it returns the path to the dataset
- The dataset is only available after installation of the package

---
class: slide

### Deal with datasets during development

- Datasets in `data/` or `inst/` are only available after package installation
- Simulate installation using `pkgload::load_all()` during development
  + `system.file()` temporarily returns development path (not installed path)

--

*Steps*

- Check your reproducible examples in the Rmd file with package data

```{r, eval=FALSE}
# For development only
pkgload::load_all()

# Same code to add in your `examples` or `tests`
# Can be tested directly during development
the_data_path <- system.file("my_dataset.csv", package = "mypackage")
the_data <- readr::read_csv(the_data_path)
```

--

> Note that `pkgload::load_all()` also loads functions in development, as it simulates a real installation

???


---
class: slide 

### Quizz

If I want to provide a dataset to the user in any format I want, where do I store it?

- A: `extdata/`
- B: `inst/`
- C: `data/`
- D: `data-raw/`


---
class: slide 

### Your turn!

- Look at the instructor demonstration with *Steps*
- Do it yourself

*Steps*  

- Create a new project
- Run `fusen::add_dev_history()`
- Run the content of the `description` chunk
- Uncomment `development` chunk in "Read data" section
- Play with data in `inst/` during development

*Bonus*

- Create function `check_data_integrity()` that reads a dataset like `nyc_squirrels` and check its integrity
  + For instance, `primary_fur_color` columns should only contains a unique color, there should not be any `+` sign inside this column.
  + Stop the function if the integrity is not good
  + Return a message if everything is ok
- Create a reproducible example using the dataset saved in `inst` and `system.file()`
- Create unit tests with reproducible examples where function should fail

???

https://github.com/statnmap/squirrels.fusen/blob/main/dev/dev_history.Rmd

```{r function-1, echo=TRUE, eval=FALSE}
#' Check data integrity
#'
#' @param x dataframe with at least columns "lat", "long" and "primary_fur_color"
#'
#' @return Original dataframe if all tests are good. Otherwise stops.
#' @export
check_data_integrity <- function(x) {
  # Verify points are in New York around Central Park
  all_coords_ok <- all(
    c(
      min(x[["lat"]]) > 40.76400,
      max(x[["lat"]]) < 40.80100,
      min(x[["long"]]) > -73.98300,
      max(x[["long"]]) < -73.94735
    )
  )
  if (!all_coords_ok) {stop("Not all data are in Central Park")}
  
  # Verify there is only one color in primary_fur_color.
  # A `+` in the column is a sign of multiple colours
  if (any(grepl("+", x[["primary_fur_color"]], fixed = TRUE))) {
    stop("There are multiple colors in some 'primary_fur_color'")
  }
  
  message("All tests are good !")
}
```

```{r examples-1, echo=TRUE, eval=FALSE}
# A working example
my_data_example <- data.frame(
  lat = c(40.77, 40.78),
  long = c(-73.95, -73.96),
  primary_fur_color = c("grey", "black")
)
check_data_integrity(my_data_example)
```
