---
title: "Versioning a {fusen} package with git"
subtitle: "And the particular case of {fusen}?"
code: "M13S02C13"
output: formation::chapitre
prerequis: [""]
obj_ped: "Know how to version a {fusen} package"
slug: versionner-fusen-git-en
---
class: slide 

## Version a {fusen} package

2 possibilities:

- Your {fusen} package has not been created yet, and you want to set up the versioning before starting the developments

- Your {fusen} package has already been created, and you now want to version it

---
class: slide

## Versioning a {fusen} package

#### Case of a new {fusen} package

- Initialize the empty project on GitLab _(remembering to use the name of your future package as project name)_

- Get the `https` link to your project by clicking on the `clone` button

- In RStudio, create the new project File > New Project > Version Control > git and link to the Repository URL

Initiate your new {fusen} package:

```{r, eval=FALSE}
fusen::create_fusen(path = ".", 
                    template = "minimal", 
                    overwrite = TRUE)
```

---
class: slide

## Versioning a {fusen} package

#### Case of a new {fusen} package

In the  *Terminal*, 

 + Switch in a new _main_ branch:
 
```{bash eval=FALSE}
git switch -c main
```

+ Select modified files in your project:

```{bash eval=FALSE}
git add .
```

+ Write a commit (with an explicit message):

```{bash eval=FALSE}
git commit -m "Init fusen package"
```

+ Push your changes back to the remote:

```{bash eval=FALSE}
git push -u origin main
```

You can now begin your developments.

---
class: slide

## Versioning a {fusen} package

#### Case of an existing {fusen} package

> This procedure can be used for any type of R project, and therefore for any type of package (not only {fusen} packages)

--

- Initialize the empty project on GitLab _(remembering to use the name of your package as project name)_

- Retrieve the `https` link to your project by clicking on the `clone` button

- Open the RStudio project and run the command `usethis::use_git()` in the console _(the {usethis} package must be installed)_

- Answer yes to all the questions asked in the console, if relevant. A first "Initial Commit" is done for you.

- RStudio restarts and git is operational locally

???

ATTENTION
Si vous modifiez cette slide, modifiez aussi M14S01C06 pour git1jour

---
class: slide

## Versioning a {fusen} package

#### Case of an existing {fusen} package

Now you have to link this project to your **`remote`**:

```{r eval=FALSE}
usethis::use_git_remote("origin", 
                        url = "https://gitlab.com/my_name/mypackage.git", 
                        overwrite = TRUE)
```

In the *Terminal*, type:

```
git push -u origin main
```

You just added a **`remote`** and made a first **`push`** on the **`main`** branch.

???

ATTENTION
Si vous modifiez cette slide, modifiez aussi M14S01C06 pour git1jour

---
class: slide

## Exercise

Set up the _git_ versioning for your `{hello}` package
