---
title: "What about data analyses in a package?"
subtitle: "From Rmd to package to Rmd"
code: ""
prerequis: [""]
output: formation::chapitre
obj_pred: ["use packages for data analyses"]
slug: data-analyses-packages
---
class: slide

## Create shareable reports through packages

### Create a package with your functions

Then, two possible ways to build your analysis reports:

- Classical package: Use external projects to use your package functions in Rmd reports
- Compendium-like: Include the Rmd reports inside the package development project

---
class: slide 

## A Compendium is like a package

The compendium logic is to separate code from the report output:

- One place to store your R scripts in "R/"
- One place to store the Rmarkdown report in "analyses/"

This is similar to package logic:

- One place to store R scripts in "R/"
- One place to store Rmarkdown examples of use in "vignettes/"

### Let's combine both worlds to benefit from robustness of packages

---
class: slide

## Combine {fusen} and Compendium structure

- Create a {fusen} project, versioned with git  

```{r, eval=FALSE}
# install.packages("fusen")
fusen::create_fusen(template = "full", with_git = TRUE)
```

- Add a "reports/" directory that will not interfer with the package

```{r, eval=FALSE}
dir.create("reports")
usethis::use_build_ignore("reports")
```


_Note: you can add the Compendium structure in the "reports/" sub-directory, with package {rcompendium}_
  
```{r, eval=FALSE}
# install.packages("rcompendium")
rcompendium::add_compendium("reports")
```

---
class: slide

## Combine {fusen} and Compendium structure

- Add a reduced dataset for the package in "inst/" or "data/"
  + Use it for examples, unit tests, documentation
  
  
- Build your functions in a "flat" file, like any {fusen} package
  + Write the content of the vignette as a reduced version of your future report
  + Create examples, unit tests, documentation as usual
  + Commit regularly
  
  
- Use the "reports/" directory to store your reports and outputs, while using the functions of your package
  + Create a Rmarkdown report using your functions loaded with `library(my.package)`
  + Do not forget to install your package prior to building your report

---
class: slide

## Why would I include my analysis inside my package?

**Because of:**

- {rim}: manage R versions
- {renv}: manage R packages versions
- and Docker: manage OS and system dependencies

---
class: slide

## Why would I include my analysis inside my package?

- Develop your project in a controled environment, with fixed versions of packages
  - Functions, tests and examples are valid for a specific set of packages versions with a specific R version
  - Use your functions for your analysis reports with the same working environment
  - Use `attachment::create_renv_for_dev()` (>=0.2.5) to build your "renv.lock" file
  
- Allow other users to create their analyses in the same environment, by cloning your repository
  + Share your work with the world, without bothering about their R installation
  
- You can build a Docker container to also fix system dependencies when using your package. {dockerfiler} can help you to set up the container.

---
class: slide

## Share your work

- Build and publish your {pkgdown} website to present how to use your package

```{r, eval=FALSE}
# Have a proper Readme - Fill and knit
usethis::use_readme_rmd()

# Allow {pkgdown}
usethis::use_pkgdown()

# Try it locally
pkgdown::build_site()

# GitHub
# Add your credentials for GitHub
gitcreds::gitcreds_set()
# Send your project to a new GitHub project
usethis::use_github()

# Build and publish with GitHub Actions
usethis::use_github_action("pkgdown")

# Build and publish on GitLab
gitlabr::use_gitlab_ci()
```

